### 2018-01-22 derivations.tex
fixes

From the ideal gas law, the number of moles of the final species is determined as:
\begin{align}
n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}, \label{e:source_moles}\\
n_{\ns} &= \frac{V P}{T \Ru} - \sum_{i=1}^{\ns - 1}{n_i}
\end{align}

\subsection{Thermo-chemical source terms}
The evolution of the thermo-chemical state variables of this system is described by a set of ordinary-differential equations:
\begin{subequations}
\label{e:source_terms}
\begin{align}
f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconp, \\
f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\}\dconv
\end{align}
\end{subequations}
For both, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
\begin{equation}
\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \label{e:spec}
\end{equation}
where $\dot{\omega}_k$ is the $kth$ species overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
\begin{subequations}
\label{e:temperature_incomplete}
\begin{align}
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}} & \text{\dconp} \\
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}} & \text{\dconv}
\end{align}
\end{subequations}
where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure specific heat and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
\begin{equation}
 [C]_{k} = \frac{n_{k}}{V}
\end{equation}

From the ideal gas law, source terms for the volume and pressure variables may derived:
\begin{subequations}
\label{e:param_incomplete}
\begin{align}
\frac{\text{d} V }{\text{d} t } &= \frac{\Ru}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconp, \\
\frac{\text{d} P }{\text{d} t } &= \frac{\Ru}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right)\dconv
\end{align}
\end{subequations}

To determine $\frac{\text{d} n }{\text{d} t }$, conservation of mass is invoked:
\begin{equation}
 \label{e:mass_conservation}
 \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n }{\text{d} t }_{k},
\end{equation}
where $W_{k}$ is the molecular weight of the $k$th species.
From~\cref{e:mass_conservation}, the source term of the last species may be written in terms of the species included in the state vector:
\begin{equation}
 \frac{\text{d} n }{\text{d} t }_{\ns} = - \frac{1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n }{\text{d} t }_{k}
 \label{e:spec_ns}
\end{equation}
Using~\cref{e:spec_ns}, the total molar rate of change can be determined:
\begin{align}
\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns} \frac{\text{d} n }{\text{d} t }_{k}, \nonumber \\
\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n }{\text{d} t }_{k}
\label{e:total_molar}
\end{align}

@@

Using~\cref{e:last_spec_conc,e:spec,e:spec_ns}, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
\begin{subequations}
\label{e:temperature_complete}
\begin{align}
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{\ns  - 1} \left(- {C_{p,\ns}} + {C_{p, k}}\right) [C]_{k}} &\text{\dconp}\\
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{\ns  - 1} \left(- {C_{v,\ns}} + {C_{v, k}}\right) [C]_{k}} &\text{\dconv}
\end{align}
\end{subequations}
This form is used for differentiation, but~\cref{e:temperature_incomplete} will often be used as well due to its compactness.

@@

\subsection{Reaction rate expressions}
The net species production rate for species $k$ is defined as:
\begin{equation}
 \label{e:spec_rop}
 \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k,i} q_{i}
\end{equation}
where $\nr$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
\begin{align}
\nu_{k,i} &= \nu^{\prime\prime}_{k,i} - \nu^{\prime}_{k,i} \\
q_{i} &= R_{i} c_{i}
\end{align}
with $\nu^{\prime}_{k,i}$ and $\nu^{\prime\prime}_{k,i}$ the product and reactant stoichiometric cofficients (respectively) of species $k$ in reaction $i$.
The base rate-of-progress for the $i$th reversible reaction $R_{i}$ is given by:
\begin{align}
R_{i} &= {R_f}_{i} - {R_r}_{i} \label{e:ropnet}\\
{R_f}_{i} &= {k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
{R_r}_{i} &= {k_r}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
\end{align}
where ${k_f}_{i}$ and ${k_r}_{i}$ are the forward and reverse reaction coefficients (respectively) for the $i$th reaction, and the third-body\slash pressure modification $c_{i}$ is given by:
\begin{equation}
\label{e:rxn_pressure}
c_i = \begin{dcases}
  \rhs{pmod}{1}[c] &\text{for elementary reactions,} \\
  \rhs{pmod}{[X]_i}[c] &\text{for third-body enhanced reactions,} \\
  \rhs{pmod}{\frac{P_{r,i}}{1 + P_{r,i}} F_i} &\text{for unimolecular\slash~recombination fall-off reactions, and} \\
  \rhs{pmod}{\frac{1}{1 + P_{r,i}} F_i} &\text{for chemically-activated bimolecular reactions,}
  \end{dcases}
\end{equation}
where for the $i$th reaction $[X]_i$ is the third-body concentration, $P_{r,i}$ is the reduced pressure, and $F_i$ is the falloff blending factor.
These terms are defined in the following sections.

The forward reaction rate coefficient $k_{f, i}$ is given by the three-parameter Arrhenius expression:
\begin{equation}
  \label{e:arrhenius}
  {k_f}_{i} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru}}\right) \;,
\end{equation}
where $A_i$ is the pre-exponential factor, $\beta_i$ is the temperature exponent, and $T_{a, i}$ is the activation temperature given by $T_{a, i} = E_{a, i} / \mathcal{R}$.

@@

\subsection{Reverse rate coefficient}
By definition, the reverse rate coefficient---${k_r}_{i}$---of irreversible reactions is zero, while reversible reactions may have non-zero ${k_r}_{i}$.
Note that in \texttt{pyJac}, reversible reactions with explicit reverse Arrhenius parameters are split into two irreversible reactions; this simplifies calculation inside the generated code, and eases comparison to Cantera~\cite{Goodwin:2015aa} which applies the same transformation.
For reversible reactions without an explicit parameterization, the reverse rate coefficient is calculated from ratio of the forward rate coefficient and the equilibrium constant:
\begin{align}
 {k_r}_{i} &= \frac{{k_f}_{i}}{{K_c}_{i}}\; , \label{e:kr}\\
 {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; ,\text{ and} \label{e:kc}\\
 {K_p}_{i} &= \text{exp}\left(\frac{\Delta S^{\circ}_k}{\Ru} - \frac{\Delta H^{\circ}_k}{\Ru T}\right) = \text{exp}\left(\sum_{k=1}^{\ns}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T}\right)\right) \label{e:kp}
\end{align}
where $P_{atm}$ is the pressure of one standard atmosphere in appropriate units.

@@

In addition to the falloff approach given previously, two additional formulations can be used to describe the pressure dependence of reactions that do not follow the modification factor $c_i$ approach.
The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa} (often termed P-Log reactions), each evaluated using~\cref{e:arrhenius}:
\begin{align}
k_1 (T) &= A_1 T^{\beta_1} \exp \left( -\frac{T_{a, 1}}{T} \right) \text{ at } P_1 \text{ and} \label{e:plog_k1} \\
k_2 (T) &= A_2 T^{\beta_2} \exp \left( -\frac{T_{a, 2}}{T} \right) \text{ at } P_2 \;, \label{e:plog_k2}
\end{align}
where the Arrhenius coefficients are given for each pressure $p_1$ and $p_2$.
Then, the reaction rate coefficient at a particular pressure $p$ between $p_1$ and $p_2$ can be determined through logarithmic interpolation:
\begin{equation}
\label{e:kf_plog}
\log{\left ({k_f}_{i} \right )} = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \log{\left (k_{1} \right )}
\end{equation}

@@

\subsection{Rate of progress derivatives}
\subsubsection{Temperature derivative}
\label{s:dri_dt}
Next, the rate of progress derivatives are evaluated, first with respect to temperature:
\begin{equation}
 \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_f}_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
\end{equation}
The derivative of the forward reaction rate coefficient is:
\begin{equation}
 \label{e:dkf_dt}
 \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
\end{equation}
While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, production rate, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
\begin{equation}
 {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
\end{equation}

Hence, the full derivative of the forward rate of progress is:
\begin{equation}
 \label{e:dropfdt_1}
 \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
\end{equation}
By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns\right\}$):
\begin{equation}
 \label{e:s_temp}
 S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
\end{equation}
and using~\cref{e:dkf_dt}, \cref{e:dropfdt_1} can be simplified to:
\begin{equation}
 \label{e:dropf_dt}
 \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_f}_{i} + \frac{{R_f}_{i}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
\end{equation}

Starting from~\cref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
\begin{equation}
 \label{e:droprdt_1}
 \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_f}_{i} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
\end{equation}
Next, \cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
\begin{align}
 \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_f}_{i} }{\text{d} T } \right) \nonumber \\
				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i}
\end{align}
The temperature derivative of the equilibrium constant is:
\begin{equation}
 \frac{\text{d} {K_c}_{i} }{\text{d} T } = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
\end{equation}
with:
\begin{equation}
 \frac{\text{d} B }{\text{d} T }_{k} = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
\end{equation}
giving:
\begin{equation}
\label{e:dkr_dt}
\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_r}_{i} \;,
\end{equation}
and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
\begin{equation}
\label{e:dropr_dt}
\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_r}_{i} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_r}_{i}
\end{equation}

>>>>

From the ideal gas law, the number of moles of the final species is determined as:
\begin{align}
n &= \frac{V P}{T \Ru} = \sum_{i=1}^{\ns}{n_i}\; , \nonumber \\
n_{\ns} &= \frac{V P}{T \Ru} - \sum_{i=1}^{\ns - 1}{n_i}
\end{align}

\subsection{Thermo-chemical source terms}
The evolution of the thermo-chemical state variables of this system is described by a set of ordinary-differential equations:
\begin{subequations}
\label{e:source_terms}
\begin{align}
f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} V }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\} & \text{\dconp} \\
f &= \frac{\text{d} \Phi }{\text{d} t } = \left\{\frac{\text{d} T }{\text{d} t },\frac{\text{d} P }{\text{d} t },\frac{\text{d} n_1}{\text{d} t },\frac{\text{d} n_2 }{\text{d} t }\ldots \frac{\text{d} n_{\ns - 1} }{\text{d} t }\right\} & \text{\dconv}
\end{align}
\end{subequations}
For both, the molar source terms are~\cite{TurnsStephenR2012Aitc}:
\begin{equation}
\frac{\text{d} n }{\text{d} t }_{k} = V \dot{\omega}_{k} \qquad k = 1, \ldots, \ns - 1, \label{e:spec}
\end{equation}
where $\dot{\omega}_k$ is the $kth$ species overall production rate, and the temperature production rate~\cite{TurnsStephenR2012Aitc} is:
\begin{subequations}
\label{e:temperature_incomplete}
\begin{align}
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} H_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{p, k}}} & \text{\dconp} \\
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns} U_{k} \dot{\omega}_{k}}{\sum_{k=1}^{\ns} [C]_{k} {C_{v, k}}} & \text{\dconv}
\end{align}
\end{subequations}
where $H_k$, $U_k$, $C_{p,k}$ and $C_{v, k}$ are the enthalpy, internal energy, constant-pressure and constant-volume specific heat of species $k$ in molar units respectively, while $[C]_{k}$ is the concentration, given by:
\begin{equation}
 [C]_{k} = \frac{n_{k}}{V}
\end{equation}

From the ideal gas law, source terms for the volume and pressure variables may derived:
\begin{subequations}
\label{e:param_incomplete}
\begin{align}
\frac{\text{d} V }{\text{d} t } &= \frac{\Ru}{P} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconp} \\
\frac{\text{d} P }{\text{d} t } &= \frac{\Ru}{V} \left(T \frac{\text{d} n }{\text{d} t } + \frac{\text{d} T }{\text{d} t } n\right) & \text{\dconv}
\end{align}
\end{subequations}

To determine $\frac{\text{d} n }{\text{d} t }$, conservation of mass is invoked:
\begin{equation}
 \label{e:mass_conservation}
 \frac{\text{d} m }{\text{d} t } = 0 = \sum_{k=1}^{\ns}  W_{k} \frac{\text{d} n_{k} }{\text{d} t },
\end{equation}
where $W_{k}$ is the molecular weight of the $k$th species.
From~\cref{e:mass_conservation}, the source term of the last species may be written in terms of the species included in the state vector:
\begin{equation}
 \frac{\text{d} n }{\text{d} t }_{\ns} = \frac{-1}{W_{\ns}} \sum_{k=1}^{\ns - 1} W_{k} \frac{\text{d} n_{k} }{\text{d} t }
 \label{e:spec_ns}
\end{equation}
Using~\cref{e:spec_ns}, the total molar rate of change can be determined:
\begin{align}
\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns} \frac{\text{d} n_{k} }{\text{d} t }, \nonumber \\
\frac{\text{d} n }{\text{d} t } &= \sum_{k=1}^{\ns - 1} \left(1 - \frac{W_{k}}{W_{\ns}}\right) \frac{\text{d} n_{k} }{\text{d} t }
\label{e:total_molar}
\end{align}

@@

Using~\cref{e:last_spec_conc,e:spec,e:spec_ns}, the temperature source terms may be expanded to contain only state variables (i.e., by removal of the last species' source term and overall production rate):
\begin{subequations}
\label{e:temperature_complete}
\begin{align}
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(H_{k} - \frac{W_{k} H_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{p,\ns}} + \sum_{k=1}^{\ns  - 1} \left({C_{p, k}} - {C_{p,\ns}}\right) [C]_{k}} &\text{\dconp}\\
\frac{\text{d} T }{\text{d} t } &= - \frac{\sum_{k=1}^{\ns  - 1} \left(U_{k} - \frac{W_{k} U_{\ns}}{W_{\ns}}\right) \dot{\omega}_{k}}{[C] {C_{v,\ns}} + \sum_{k=1}^{\ns  - 1} \left({C_{v, k}} - {C_{v,\ns}}\right) [C]_{k}} &\text{\dconv}
\end{align}
\end{subequations}
This form of the temperature source term will be used for derivation of Jacobian terms, but~\cref{e:temperature_incomplete} will often be used as well due to its compactness.

@@

\subsection{Reaction rate expressions}
The net species production rate for species $k$ is defined as:
\begin{equation}
 \label{e:spec_rop}
 \dot{\omega}_{k} = \sum_{i=1}^{\nr} \nu_{k,i} q_{i}
\end{equation}
where $\nr$ is the number of reactions in the chemical kinetic model, $\nu_{k, i}$ the overall stoichiometric coefficient of species $k$ in reaction $i$ and $q_i$ the net rate-of-progress for reaction $i$:
\begin{align}
\nu_{k,i} &= \nu^{\prime\prime}_{k,i} - \nu^{\prime}_{k,i} \\
q_{i} &= R_{i} c_{i}
\end{align}
with $\nu^{\prime}_{k,i}$ and $\nu^{\prime\prime}_{k,i}$ the product and reactant stoichiometric cofficients (respectively) of species $k$ in reaction $i$.
The base rate-of-progress for the $i$th reversible reaction $R_{i}$ is given by:
\begin{align}
R_{i} &= {R_{f,i}} - {R_{r,i}} \label{e:ropnet}\\
{R_{f, i}} &= {k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} \label{e:ropf}\\
{R_{r, i}} &= {k_{r, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} \label{e:ropr}
\end{align}
where ${k_{f, i}}$ and ${k_{r, i}}$ are the forward and reverse reaction coefficients (respectively) for the $i$th reaction, and the third-body\slash pressure modification $c_{i}$ is given by:
\begin{equation}
\label{e:rxn_pressure}
c_i = \begin{dcases}
  \rhs{pmod}{1}[c] &\text{for elementary reactions,} \\
  \rhs{pmod}{[X]_i}[c] &\text{for third-body enhanced reactions,} \\
  \rhs{pmod}{\frac{P_{r,i}}{1 + P_{r,i}} F_i} &\text{for unimolecular\slash~recombination fall-off reactions, and} \\
  \rhs{pmod}{\frac{1}{1 + P_{r,i}} F_i} &\text{for chemically-activated bimolecular reactions,}
  \end{dcases}
\end{equation}
where for the $i$th reaction $[X]_i$ is the third-body concentration, $P_{r,i}$ is the reduced pressure, and $F_i$ is the falloff blending factor.
These terms are defined in the following sections.

The forward reaction rate coefficient $k_{f, i}$ is given by the three-parameter Arrhenius expression:
\begin{equation}
  \label{e:arrhenius}
  {k_{f, i}} = A_{i} T^{\beta_{i}} \operatorname{exp}\left({- \frac{{E_{a}}_{i}}{T \Ru}}\right) \;,
\end{equation}
where $A_i$ is the pre-exponential factor, $\beta_i$ is the temperature exponent, and $T_{a, i}$ is the activation temperature given by $T_{a, i} = E_{a, i} / \mathcal{R}$.

@@

\subsection{Reverse rate coefficient}
By definition, the reverse rate coefficient---${k_{r, i}}$---of irreversible reactions is zero, while reversible reactions may have non-zero ${k_{r, i}}$.
Note that in \texttt{pyJac}, reversible reactions with explicit reverse Arrhenius parameters are split into two irreversible reactions; this simplifies calculation inside the generated code, and eases comparison to Cantera~\cite{Goodwin:2015aa} which applies the same transformation.
For reversible reactions without an explicit parameterization, the reverse rate coefficient is calculated from ratio of the forward rate coefficient and the equilibrium constant:
\begin{align}
 {k_{r, i}} &= \frac{{k_{f, i}}}{{K_c}_{i}}\; , \label{e:kr}\\
 {K_c}_{i} &= \left(\left(\frac{P_{atm}}{T \Ru}\right)^{\sum_{k=1}^{\ns} \nu_{k,i}}\right) {K_p}_{i}\; ,\text{ and} \label{e:kc}\\
 {K_p}_{i} &= \text{exp}\left(\frac{\Delta S^{\circ}_k}{\Ru} - \frac{\Delta H^{\circ}_k}{\Ru T}\right) = \text{exp}\left(\sum_{k=1}^{\ns}\nu_{ki}\left(\frac{S^{\circ}_k}{\Ru} - \frac{H^{\circ}_k}{\Ru T}\right)\right) \label{e:kp}
\end{align}
where $P_{atm}$ is the pressure of one standard atmosphere in appropriate units.

@@

In addition to the falloff approach given previously, two additional formulations can be used to describe the pressure dependence of reactions that do not follow the modification factor $c_i$ approach.
The first involves logarithmic interpolation between Arrhenius rates at two pressures~\cite{chemkin:2012,Goodwin:2015aa} (often termed P-Log reactions), each evaluated using~\cref{e:arrhenius}:
\begin{align}
k_1 (T) &= A_1 T^{\beta_1} \exp \left( -\frac{T_{a, 1}}{T} \right) \text{ at } P_1 \text{ and} \label{e:plog_k1} \\
k_2 (T) &= A_2 T^{\beta_2} \exp \left( -\frac{T_{a, 2}}{T} \right) \text{ at } P_2 \;, \label{e:plog_k2}
\end{align}
where the Arrhenius coefficients are given for each pressure $p_1$ and $p_2$.
Then, the reaction rate coefficient at a particular pressure $p$ between $p_1$ and $p_2$ can be determined through logarithmic interpolation:
\begin{equation}
\label{e:kf_plog}
\log{\left ({k_{f, i}} \right )} = \frac{\left(- \log{\left (k_{1} \right )} + \log{\left (k_{2} \right )}\right) \left(- \log{\left (P_{1} \right )} + \log{\left (P \right )}\right)}{- \log{\left (P_{1} \right )} + \log{\left (P_{2} \right )}} + \log{\left (k_{1} \right )}
\end{equation}

@@

\subsection{Rate of progress derivatives}
\subsubsection{Temperature derivative}
\label{s:dri_dt}
Next, the rate of progress derivatives are evaluated, first with respect to temperature:
\begin{equation}
 \frac{\partial R }{\partial T }_{i} = \frac{\partial}{\partial T}\left({k_{f, i}} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}}\right)
\end{equation}
The derivative of the forward reaction rate coefficient is:
\begin{equation}
 \label{e:dkf_dt}
 \frac{\text{d} {k_f} }{\text{d} T }_{i} = \frac{{k_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
\end{equation}
While evaluating derivatives for the Jacobian, it is important to expand terms---e.g., concentration, production rate, etc.---related to the last species to obtain the correct derivative, in this case using \cref{e:last_spec_conc}:
\begin{equation}
 {R_f} = \left(\left(- \sum_{k=1}^{\ns  - 1} [C]_{k} + \frac{P}{T \Ru}\right)^{\nu^{\prime}_{\ns,i}}\right) {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
\end{equation}

Hence, the full derivative of the forward rate of progress is:
\begin{equation}
 \label{e:dropfdt_1}
 \frac{\partial {R_f} }{\partial T }_{i} = \frac{\text{d} {k_f} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime}_{k,i}} - \frac{[C] \nu^{\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime}_{k,i}}
\end{equation}
By defining a temporary variable $S^{\prime}_{l}$ (for species $l = \left\{j, \ns\right\}$):
\begin{equation}
 \label{e:s_temp}
 S^{\prime}_{l} = \nu^{\prime}_{l,i} [C]_{l}^{\nu^{\prime}_{l,i} - 1} \prod_{\substack{1 \leq l \leq l - 1\\l + 1 \leq l \leq \ns}} [C]_{l}^{\nu^{\prime}_{l,i}}
\end{equation}
and using~\cref{e:dkf_dt}, \cref{e:dropfdt_1} can be simplified to:
\begin{equation}
 \label{e:dropf_dt}
 \frac{\partial {R_f} }{\partial T }_{i} = - \frac{[C] S^{\prime}_{\ns}}{T} {k_{f, i}} + \frac{{R_{f, i}}}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)
\end{equation}

Starting from~\cref{e:ropr} and applying the same expansion of the last species' concentration as previously, the temperature derivative of the reverse rate of progress is:
\begin{equation}
 \label{e:droprdt_1}
 \frac{\partial {R_r} }{\partial T }_{i} = \frac{\text{d} {k_r} }{\text{d} T }_{i} \prod_{k=1}^{\ns} [C]_{k}^{\nu^{\prime\prime}_{k,i}} - \frac{[C] \nu^{\prime\prime}_{\ns,i}}{T} [C]_{\ns}^{\nu^{\prime\prime}_{\ns,i} - 1} {k_{f, i}} \prod_{k=1}^{\ns  - 1} [C]_{k}^{\nu^{\prime\prime}_{k,i}}
\end{equation}
Next, \cref{e:kr} is considered to obtain the temperature derivative of the non-explicit reversible reaction rate coefficient:
\begin{align}
 \frac{\text{d} {k_r} }{\text{d} T }_{i} &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } \frac{{k_f}_i}{{K_c}_i} + \frac{1}{{K_c}_{i}}\frac{\text{d} {k_{f, i}} }{\text{d} T } \right) \nonumber \\
				         &= \left(- \frac{1}{{K_c}_{i}} \frac{\text{d} {K_c}_{i} }{\text{d} T } + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}}
\end{align}
The temperature derivative of the equilibrium constant is:
\begin{equation}
 \frac{\text{d} {K_c}_{i} }{\text{d} T } = {K_c}_{i} \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} \;,
\end{equation}
with:
\begin{equation}
 \frac{\text{d} B }{\text{d} T }_{k} = T \left(T \left(\frac{T a_{k,4}}{5} + \frac{a_{k,3}}{4}\right) + \frac{a_{k,2}}{3}\right) + \frac{a_{k,1}}{2} + \frac{1}{T} \left(a_{k,0} - 1 + \frac{a_{k,5}}{T}\right)
\end{equation}
giving:
\begin{equation}
\label{e:dkr_dt}
\frac{\text{d} {k_r} }{\text{d} T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {k_{r, i}} \;,
\end{equation}
and using an temporary value $S^{\prime\prime}_{\ns}$---defined analagously to \cref{e:s_temp} for the reverse direction---we obtain:
\begin{equation}
\label{e:dropr_dt}
\frac{\partial {R_r} }{\partial T }_{i} = \left(- \sum_{k=1}^{\ns} \nu_{k,i} \frac{\text{d} B }{\text{d} T }_{k} + \frac{1}{T} \left(\beta_{i} + \frac{{E_{a}}_{i}}{T \Ru}\right)\right) {R_{r, i}} - \frac{[C] S^{\prime\prime}_{\ns}}{T} {k_{r, i}}
\end{equation}
